Title

### 2024-06-27 paper/main.tex
Updates

Understanding cause-and-effect relationships between variables is a fundamental
goal in many scientific questions as it reveals the underlying mechanisms of
the observed phenomena and informs effective interventions or policy decisions.
\emph{Causal Discovery} methods aim to uncover the underlying causal structure
among random variables from a given dataset. The problem of causal discovery
has been explored in both the Directed Acyclic Graphs (DAGs) and Structural
Equation Models (SEM) literatures, each with different approaches. In the DAG
literature, the focus has been on developing automated algorithms that can
learn the causal structure from observational or interventional dataset. In
contrast, the SEM literature emphasizes tools that help researchers manually
construct or modify modify.

@@

\begin{enumerate}
	\item \textbf{Trustworthiness: } Although most algorithms have
		asymptotic theory proving them to be consistent, their finite
		sample properties are not well understood. In real-world
		settings with finite samples, these algorithms can make obvious
		mistakes, for example, the constraint-based algorithms are
		biased towards sparse graphs \citep{Malinsky2024}. Consequently,
		researchers often have rely on post-hoc modifications based on 
		their domain knowledge or model testing methods.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
		structures can be faithful to a given observational dataset.
		Thus, using only observational data, these algorithms can only
		recover the MECs that can contain unoriented edges. However,
		most methods for downstream tasks, such as identification or
		causal effect estimation, assume knowledge of a fully oriented
		DAG. As a results, researchers have to rely on their expertise
		to orient these edges or make further assumptions to make use
		of pairwise orientation techniques.
	\item \textbf{Algorithm and hyperparameter selection:} Each algorithm
		has the same goal but they can learn very different models
		depending on the choice of algorithm and their hyperparameters.
		For example, the choice of the conditional independence (CI)
		test being used for constraint-based method depends on the
		dataset, similarly for the scoring method for score-based
		methods. With no obvious way to test the impact/performance of
		these choices on a given dataset, it becomes quite hard to
		chose the optimal algorithm for a given dataset.
\end{enumerate}

As a result of these challenges, no matter if the researcher is manually
constructing the DAG or using an algorithm, they need to utilize their domain
expertise to construct and modify models. One potential way to through
conditional independence (CI) based testing. As each missing edge in the DAG
implies a CI statement, that can be tested in our data \ref{fig:ci_testing}.
This however has two main drawbacks: 
\begin{enumerate}
	\item As CIs are implied by only missing edges, we can not test if an
		incorrect edge is present in the model.
	\item For larger graphs, the number of CI tests can be very large, for example,
		the alarm network with $ 37 $ nodes has $ 287 $ implied CIs. As
		there is no way to rank these CIs, it can be tedious to
		consider all these statements.
\end{enumerate}

@@

On the other hand, in Structural Equation Models (SEMs) literature, more
emphasis has been placed on helping researchers manually construct models while
integrating their expertise. Researchers typically develop an initial model
based on domain knowledge and use various tools to guide modifications to
improve the fit of their model to data while it still making theoretical sense.
This model construction process is known as \emph{Specification Search}. A
common method for specification search is \emph{modification indices}, which
identify potential changes to the model structure that improve the model fit
and ranks them using the most improvement to the model fit. Researchers then
use these suggestions and their expertise to choose the most optimal
modifications.

>>>>

Understanding cause-and-effect relationships between variables is a fundamental
goal in many scientific questions as it reveals the underlying mechanisms of
the observed phenomena and informs effective interventions or policy decisions.
\emph{Causal Discovery} methods aim to uncover the underlying causal structure
among random variables from a given observational dataset. The problem of
causal discovery has been explored in both the Directed Acyclic Graphs (DAGs)
and Structural Equation Models (SEMs) literatures, each with different
approaches. In the DAG literature, the focus has been on developing automated
algorithms that can learn the causal structure from observational datasets. In
contrast, the SEM literature focuses on tools that help researchers manually
construct or modify models.

@@

\begin{enumerate}
	\item \textbf{Trustworthiness: } Although most algorithms have
		asymptotic theory proving them to be consistent, their finite
		sample properties are not well understood. In real-world
		settings with finite samples, these algorithms can make obvious
		mistakes, for example, the constraint-based algorithms are
		biased towards sparse graphs \citep{Malinsky2024}.
		Consequently, researchers often have rely on post-hoc
		modifications based on model testing methods and their domain
		knowledge.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
		structures can be faithful to a given observational dataset.
		Thus, using only observational data, these algorithms can only
		recover the MECs that can contain unoriented edges. However,
		most methods for downstream tasks, such as identification or
		causal effect estimation, assume knowledge of a fully oriented
		DAG. As a result, researchers have to rely on their expertise
		to orient these edges or make further assumptions to make use
		of pairwise orientation techniques.
	\item \textbf{Algorithm and hyperparameter selection:} Each algorithm
		has the same goal but they can learn very different models
		depending on the choice of algorithm and their hyperparameters.
		For example, the choice of the conditional independence (CI)
		test being used for constraint-based method depends on the
		dataset, similarly for the scoring method for score-based
		methods. With no obvious way to test the impact/performance of
		these choices on a given dataset, it becomes quite hard to
		chose the optimal algorithm for a given dataset.
\end{enumerate}

As a result of these challenges, doesn't matter if the researcher is manually
constructing the DAG or using an algorithm, they need to do some manual
modifications to the model before being able to use them for analysis. One of
the common ways to guide this modification is to use Conditional Independence
(CI) based model testing \cite{Ankan2021}. As DAGs imply CI statements that can
be tested in the given data, and then model can be modified based on the
results of this test. Figure~\ref{fig:ci_testing} shows an example of testing
using CIs. This approach has two main drawbacks:
\begin{enumerate}
	\item CIs are implied by missing edges only, and hence we can not test
		whether an incorrect edge is present in the model.
	\item For larger graphs, the number of CI tests can be very large, for
		example, the alarm network with $ 37 $ nodes has $ 287 $
		implied CIs. It can be difficult to look at all these tests for
		violations.
\end{enumerate}

@@

On the other hand, in Structural Equation Models (SEMs) literature, more
emphasis has been placed on helping researchers manually construct models while
integrating their expertise. Researchers typically develop an initial model
based on domain knowledge and use various tools to guide modifications to
improve the fit of their model to data while it still making theoretical sense.
This model construction process is known as \emph{Specification Search}. A
common method for specification search is \emph{modification indices}, which
identify potential changes to the model structure that improve the model fit
and ranks them using the most improvement to the model fit. This ranking allows
the researchers to focus first on the suggestions that can improve the model
the most and can use their expertise to choose the most optimal modifications.
